Neon: Nuclear Norm to Beat Muon

### 2025-05-07 article/muon.tex
Added mentors

\author{
  Alexey Kravatskiy\\
  \texttt{kravtskii.aiu@phystech.edu}
  \and
  Ivan Kozyrev\\
  \texttt{kozyrev.in@phystech.edu}
  \and
  Nikolay Kozlov \\
  \texttt{kozlov.na@phystech.edu}
  \and
  Alexander Vinogradov \\
  \texttt{vinogradov.am@phystech.edu}
}

@@

    \item Neon (rank-1 version), Muon, AdamW and SGD are compared on MLP and CNN (see Figures~\ref{fig:mlp_epochs}, \ref{fig:cnn_epochs}, \ref{fig:mlp_time}, and~\ref{fig:cnn_time}). All methods work correctly, but again there is the problem with which one is the fastest (for now, it's SGD).
    \begin{figure}[h!]
        \center{\includegraphics[width=0.8\linewidth]{../figs/mlp24/accuracy_vs_epochs_simple.png}}
        \caption{MLP: self.linear1 = nn.Linear(32*32*3, 512), self.linear2 = nn.Linear(512, 10), self.activ = nn.GELU()}
        \label{fig:mlp_epochs}
    \end{figure}
    \begin{figure}[h!]
        \center{\includegraphics[width=0.8\linewidth]{../figs/mlp24/accuracy_vs_epochs_moderate.png}}
        \caption{CNN: 2 convolutional blocks, 2 fully connected layers, activation + dropout}
        \label{fig:cnn_epochs}
    \end{figure}
    \begin{figure}[h!]
        \center{\includegraphics[width=0.8\linewidth]{../figs/mlp24/accuracy_vs_time_simple.png}}
        \caption{MLP: wallclock time measurements}
        \label{fig:mlp_time}
    \end{figure}
    \begin{figure}[h!]
        \center{\includegraphics[width=0.8\linewidth]{../figs/mlp24/accuracy_vs_time_moderate.png}}
        \caption{CNN: wallclock time measurements}
        \label{fig:cnn_time}
    \end{figure}
    \begin{table}[h!]
        \centering
        \begin{tabular}{c|c|c|c}
        \hline
        Method                              & rtol        & k & time,s \\
        \hline 
        Power Iterations                    & 0.01        & 1 & 7.7\\ 
        SVDS (thick-restart Lanczos method) & 0.01        & 1 & 0.18\\
        PCA Low Rank (RSVD)                 & 0.01        & 1 & 1.15\\
        SVDS (thick-restart Lanczos method) & 0.01        & 10 & 0.47\\
        PCA Low Rank (RSVD)                 & 0.01        & 10 & 19.4\\
        SVDS (thick-restart Lanczos method) & 0.01        & 100 & 1.96\\
        PCA Low Rank (RSVD)                 & 0.01        & 100 & 170\\
        \end{tabular}
        \caption{k-rank updated comparison}
        Comparison of different numerical methods to calculate k-rank update on $5000\times5000$ matrix of real numbers, rtol is an error in Frobenius norm relative to the k-rank approximation of truncated svd. During the research it was noted that rsvd can give good and fast approximation for singular values, but the matrix of approximation is far from the one given by truncated svd, while Lanczos method gives good and fast approximation for a matrix, but not so good approximation for singular values. 
        \label{tab:matrix_methods}
    \end{table}
\end{enumerate}

>>>>

\author{
  Alexey Kravatskiy\\
  \texttt{kravtskii.aiu@phystech.edu}
  \and
  Ivan Kozyrev\\
  \texttt{kozyrev.in@phystech.edu}
  \and
  Nikolay Kozlov \\
  \texttt{kozlov.na@phystech.edu}
  \and
  Alexander Vinogradov \\
  \texttt{vinogradov.am@phystech.edu}
  \and
  Daniil Merkulov \\
  \texttt{daniil.merkulov@phystech.edu}
  \and
  Ivan Oseledets \\
  \texttt{i.oseledets@scholtech.ru}
}

@@

    \item Neon (rank-1 version), Muon, AdamW and SGD are compared on MLP and CNN (see Figures~\ref{fig:mlp_epochs}, \ref{fig:cnn_epochs}, \ref{fig:mlp_time}, and~\ref{fig:cnn_time}). All methods work correctly, but again there is the problem with which one is the fastest (for now, it's SGD).
    \begin{figure}[h!]
        \center{\includegraphics[width=0.8\linewidth]{../figs/mlp0605/loss_vs_epochs_mlp.pdf}}
        \caption{MLP: self.linear1 = nn.Linear(32*32*3, 512), self.linear2 = nn.Linear(512, 10), self.activ = nn.GELU()}
        \label{fig:mlp_epochs}
    \end{figure}
    \begin{figure}[h!]
        \center{\includegraphics[width=0.8\linewidth]{../figs/mlp24/accuracy_vs_epochs_moderate.png}}
        \caption{CNN: 2 convolutional blocks, 2 fully connected layers, activation + dropout}
        \label{fig:cnn_epochs}
    \end{figure}
    \begin{figure}[h!]
        \center{\includegraphics[width=0.8\linewidth]{../figs/mlp0605/accuracy_vs_time_mlp.pdf}}
        \caption{MLP: wallclock time measurements}
        \label{fig:mlp_time}
    \end{figure}
    \begin{figure}[h!]
        \center{\includegraphics[width=0.8\linewidth]{../figs/mlp24/accuracy_vs_time_moderate.png}}
        \caption{CNN: wallclock time measurements}
        \label{fig:cnn_time}
    \end{figure}
    \begin{table}[h!]
        \centering
        \begin{tabular}{c|c|c|c}
        \hline
        Method                              & rtol        & k & time,s \\
        \hline 
        Power Iterations                    & 0.01        & 1 & 7.7\\ 
        SVDS (thick-restart Lanczos method) & 0.01        & 1 & 0.18\\
        PCA Low Rank (RSVD)                 & 0.01        & 1 & 1.15\\
        SVDS (thick-restart Lanczos method) & 0.01        & 10 & 0.47\\
        PCA Low Rank (RSVD)                 & 0.01        & 10 & 19.4\\
        SVDS (thick-restart Lanczos method) & 0.01        & 100 & 1.96\\
        PCA Low Rank (RSVD)                 & 0.01        & 100 & 170\\
        \end{tabular}
        \caption{k-rank updated comparison}
        Comparison of different numerical methods to calculate k-rank update on $5000\times5000$ matrix of real numbers, rtol is an error in Frobenius norm relative to the k-rank approximation of truncated svd. During the research it was noted that rsvd can give good and fast approximation for singular values, but the matrix of approximation is far from the one given by truncated svd, while Lanczos method gives good and fast approximation for a matrix, but not so good approximation for singular values. 
        \label{tab:matrix_methods}
    \end{table}
\end{enumerate}
